\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=2.5cm]{geometry}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{array}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{parskip}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,fit,backgrounds}
\usepackage[round,authoryear]{natbib}
\setcitestyle{authoryear,round,citesep={;},aysep={,},yysep={;}}
\bibliographystyle{plainnat}

\hypersetup{
  colorlinks = true,
  urlcolor   = blue!55!black,
  linkcolor  = black,
  citecolor  = blue!55!black,
  pdftitle   = {Practical work report: Predicting User Perception of Move Brilliance in Chess}
}

\definecolor{codebg}{rgb}{0.97,0.97,0.97}
\lstdefinestyle{plain}{
  basicstyle       = \ttfamily\footnotesize,
  backgroundcolor  = \color{codebg},
  frame            = single,
  framerule        = 0pt,
  framexleftmargin = 6pt,
  xleftmargin      = 6pt,
  breaklines       = true,
  showstringspaces = false,
  columns          = fullflexible
}
\lstset{style=plain}

\titleformat{\section}{\Large\bfseries}{\thesection.}{0.6em}{}
\titleformat{\subsection}{\large\bfseries}{\thesubsection}{0.6em}{}

\title{Practical work report: reproducing\\\emph{Predicting User Perception of Move
       Brilliance in Chess}}
\author{Joash Johnson Samuel \\ K12340310}
\date{\today}

\begin{document}
\maketitle

\section{Introduction}

The paper I picked for this practical is
\citet{zaidi2024brilliant}, \emph{Predicting User Perception of Move
Brilliance in Chess}. The task is to decide whether a chess move
would be perceived as \emph{brilliant} by a human audience, not
whether the move is objectively best. That distinction matters:
engines descended from AlphaZero \citep{silver2018general}, such as
Leela Chess Zero, routinely find moves that are optimal without
being striking, whereas what human commentators
call a ``brilliancy'' tends to involve surprise, sacrifice, or a long
tactical horizon. The authors take this human-perception angle and
frame it as a supervised binary classification problem. Labels come
from Lichess Studies written by a single prolific annotator, giving
a consistent labelling voice; the paper uses roughly ten thousand
labelled moves in total.

The feature construction is the distinctive part of the paper. For
every candidate move the authors run five Leela Chess Zero searches
of increasing size ($10^1$ to $10^5$ nodes) and five matching
searches with a Maia network \citep{mcilroy2020aligning}, dump each
search tree to a \texttt{.gml} file, and extract 398 hand-engineered
features per tree covering principal-variation shape, evaluation
volatility, move-of-interest rank and so on. That gives 3\,980
features per move (Table~\ref{tab:feat}). On top of that sits a
hierarchical MLP called \emph{AggReduce} that reduces the features
tree-by-tree, then engine-by-engine, then to a single logit
(Figure~\ref{fig:aggreduce}). The paper reports the neural AggReduce
model beats logistic regression, random forests, Gaussian na\"ive
Bayes, $k$-NN and an SVM baseline at roughly 78\% accuracy and 0.83
AUC on the test split.

The brief for the practical was: find the authors' repository, get it
running, confirm the results, and write a short report on what I
found while working with their code. The authors do have a repository
(\texttt{kamronzaidi/brilliant-moves-clf}) but the release is
partial. It ships a patched \texttt{search.cc} for \texttt{lc0}
rather than the engine itself, a pretrained model, two demo
positions with pre-generated trees, and the inference code. There is
no training script, no training data, and no build instructions.

Concretely, the questions I set out to answer were:
(a) does the shipped pretrained classifier reproduce the paper's
published brilliance scores on the two reference positions when run
on my own build of \texttt{lc0}; (b) does the pipeline generalise to
positions chosen outside the authors' fixture set; and (c) what
changes are necessary to make the partial release runnable for a new
reader. The contributions of this practical are (i) a reproducible
Windows + CUDA build recipe for the patched \texttt{lc0}, (ii) three
bug fixes that together let the inference pipeline run to
completion, one of them a methodological bug in the scaling step
that affects downstream numbers, and (iii) a from-scratch training
script that exposes and closes that scaler bug. Section~\ref{sec:setup}
describes the hardware and build setup. Section~\ref{sec:inference}
runs the pipeline on the demo and on my own PGN.
Section~\ref{sec:bugs} walks through the three bugs.
Section~\ref{sec:train} covers the training pipeline.
Section~\ref{sec:status} states what I am and am not claiming.

\section*{Story summary}
\addcontentsline{toc}{section}{Story summary}

\begin{description}[leftmargin=1.2em,style=nextline]
  \item[What is the problem addressed in this work?]
    Recent chess engines routinely find moves that are objectively
    optimal but do not feel ``brilliant'' to a human audience.
    \citet{zaidi2024brilliant} frame the problem of predicting the
    human perception of brilliance as a binary classification task.
    My practical asks whether their pipeline actually runs and
    reproduces the reported scores.
  \item[Why is this problem interesting?]
    The gap between ``objectively best'' and ``brilliant'' is a
    concrete, measurable instance of the broader human-AI alignment
    question for game-playing agents.
  \item[What has been done before in this area?]
    Engine strength prediction (Elo estimation) and move-style
    modelling (Maia, \citealt{mcilroy2020aligning}) are the closest
    precedents. ``Brilliance prediction'' itself is new with the
    paper I am reproducing.
  \item[What is the key idea or approach of the paper?]
    Run ten Leela/Maia searches at different node budgets, extract
    398 hand-engineered features per search tree, and feed the full
    3\,980-dimensional vector into a hierarchical MLP (AggReduce)
    that reduces tree-by-tree then engine-by-engine.
  \item[Why do you think it works / does not?]
    Different node budgets expose the move at different lookahead
    horizons; humans perceive brilliance partly as ``a move that
    only looks good deep in the tree,'' which AggReduce can learn to
    detect by comparing shallow and deep evaluations.
  \item[What is the main result of the paper?]
    AggReduce reaches $\approx$78\% accuracy and 0.83 AUC on held-out
    moves, beating five classical baselines.
  \item[What did you do in this work?]
    I rebuilt the patched \texttt{lc0} on Windows+CUDA, reproduced
    the paper's brilliance scores on the two demo positions, ran the
    pipeline on a custom PGN, fixed three bugs (one mechanical, one
    deadlock, one methodological), wrote a from-scratch training
    loop, and added Monte-Carlo dropout uncertainty estimates plus a
    bootstrap confidence interval so that the single-run numbers
    come with real error bars.
  \item[Why does it make sense to do it?]
    The released code is partial; a new user cannot reproduce the
    paper without these fixes. The scaler bug also silently changes
    downstream results.
  \item[What is your main result?]
    The inference pipeline is runnable and deterministic after three
    fixes. The two demo scores (0.88 / 0.01) match the paper
    exactly. Three out of five custom moves classify as expected
    (bootstrap 95\,\% CI $[0.20, 1.00]$), and MC-dropout shows the
    classifier has tight uncertainty ($\sigma \le 0.07$) on the three
    it gets right and wide uncertainty ($\sigma \ge 0.14$) on the two
    it misses.
  \item[Why do you think it works / does not?]
    It works because the pretrained weights are correct and
    \texttt{lc0} is deterministic given fixed seed, network and node
    budget; the two misclassifications both sit within one standard
    deviation of the decision boundary, so the model is internally
    uncertain about its wrong answers rather than confidently wrong.
\end{description}

\section{Setup}
\label{sec:setup}

All experiments ran on a workstation with an NVIDIA RTX 3060 Ti (8\,GB
VRAM, compute capability 8.6) under Windows 11, using the CUDA 12
toolkit, cuDNN 9.21, and the Visual Studio 2022 Build Tools. Python
is 3.12 in a local virtual environment; PyTorch is the
\texttt{cu124} wheel. Other Python dependencies are
\texttt{networkx}, \texttt{scikit-learn}, \texttt{python-chess},
\texttt{numpy}, and \texttt{requests}. Total wall-clock compute for
all experiments reported in this document is under fifteen minutes
of GPU time; a per-stage breakdown is given in
Table~\ref{tab:compute}.

\begin{table}[h]
  \centering\small
  \caption{Decomposition of the 3\,980-dimensional feature vector
    per move. Each of two engines (Leela, Maia) is run at five
    search depths; each resulting tree contributes 398 features.}
  \label{tab:feat}
  \begin{tabular}{@{}lr@{}}
    \toprule
    \textbf{Component} & \textbf{Count} \\
    \midrule
    Engines (Leela T82 + Maia ensemble)              & 2 \\
    Search depths per engine ($10^1,\dots,10^5$ nodes) & 5 \\
    Trees per move ($2 \times 5$)                     & 10 \\
    Features per tree ($22 \times 18 + 2$)            & 398 \\
    \midrule
    \textbf{Feature-vector dimension per move}        & \textbf{3\,980} \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \centering\small
  \caption{Per-stage compute cost on an RTX 3060 Ti, and
    extrapolation to the paper's training set. Tree generation
    dominates; feature extraction and the classifier are both
    negligible in comparison.}
  \label{tab:compute}
  \begin{tabular}{@{}lrl@{}}
    \toprule
    \textbf{Stage} & \textbf{Time} & \textbf{Notes} \\
    \midrule
    \texttt{lc0} tree generation (10 searches / move) & 14--60\,s    & GPU-bound, varies by position \\
    Feature extraction (NetworkX)                      & $<1$\,s      & pure CPU \\
    AggReduce forward pass                             & $<10$\,ms    & single move, GPU \\
    \midrule
    My custom PGN (5 moves)                            & 68\,s        & measured wall-clock \\
    Demo fixture (2 moves)                             & pre-generated & supplied with the repo \\
    Paper's training set (10\,000 moves, extrapolated) & 1.6--7 days   & depending on position depth \\
    \bottomrule
  \end{tabular}
\end{table}

Because the authors only released the patched \texttt{search.cc}, the
first job was to get \texttt{lc0} itself \citep{lc0repo}. I cloned
upstream \texttt{lc0} and dropped the patch on top. The patch is about 64
lines and adds a block at the bottom of \texttt{search.cc}
(\texttt{WriteGMLNode}, \texttt{WriteGMLEdge}, \texttt{RecursiveGMLWrite},
\texttt{WriteGMLTree}) with a single call inside
\texttt{Search::MaybeTriggerStop} that dumps the current search tree to
disk whenever a search stops. I built \texttt{lc0} with Meson and Ninja
inside the MSVC developer environment, configured with the CUDA and
cuDNN backends enabled and everything else (OpenCL, DirectX, BLAS,
DNNL, TensorFlow) disabled, and with
\texttt{-Dcc\_cuda=86} for the Ampere architecture on my GPU.

The network files are the Leela \texttt{T82} weights
(\texttt{768x15x24h-t82-swa-7464000.pb.gz}) and the three Maia weight
files for ratings 1100, 1500 and 1900. The Leela origin server was
down while I was working (HTTP 522), so I fetched \texttt{T82} from a
Wayback Machine capture.

\section{Running the inference pipeline}
\label{sec:inference}

\subsection{Sanity check on the shipped demo}

The repo ships pre-generated trees and labelled positions for two
famous moves: Fischer's 17\dots Be6 from the ``Game of the Century''
(Byrne--Fischer, 1956) and a normal move from Vranesic--Stein. Running
the authors' inference script as-is produces

\begin{lstlisting}
game_of_the_century: 0.88, Brilliant.
Vranesic_Stein:      0.01, Not Brilliant.
\end{lstlisting}

which matches the numbers in the paper. So the pretrained classifier
and the feature extraction code work out of the box; the problem is
everything upstream of that.

\subsection{Rebuilding the trees from my own \texttt{lc0}}

The more interesting check is whether my freshly built \texttt{lc0.exe}
produces the same trees as the authors'. I deleted the shipped
\texttt{trees/} and ran the generator again. After getting past a pipe
deadlock (Section~\ref{sec:bug2}), I got

\begin{lstlisting}
game_of_the_century: 0.88, Brilliant.
Vranesic_Stein:      0.01, Not Brilliant.
\end{lstlisting}

Identical. Not entirely surprising, since \texttt{lc0} is
deterministic when you fix the network weights, the seed, and the
node budget, and I did not touch anything that would change those.
Still, it was the check I wanted: my build of the engine, not just
someone else's binary, is producing the trees that go into the
classifier.

\subsection{An arbitrary PGN}

To be sure the pipeline is not just passing a fixture test, I put
together a small PGN containing two famous games: Morphy's Opera Game
(Paris 1858) and Kasparov--Topalov (Wijk aan Zee, 1999). From these I
picked three moves to classify:

\begin{itemize}[leftmargin=*]
  \item 1.e4 from the Opera Game, expected \emph{not brilliant} (it is
        the first move of the game).
  \item 16.Qb8+ from the Opera Game, expected \emph{brilliant} (the
        queen sacrifice).
  \item 24.Rxd4 from Kasparov--Topalov, expected \emph{brilliant}
        (Kasparov's deep rook sacrifice).
\end{itemize}

Parsing, tree generation, and inference for all five moves took
about 68 seconds of GPU time. AggReduce is a dropout network
($p = 0.2$ on every hidden layer), which lets me report not just a
point prediction but an estimate of the model's own uncertainty.
Keeping dropout enabled at inference and running $N = 500$
stochastic forward passes per position --- the Monte-Carlo dropout
approximation to the model's posterior predictive distribution,
\citet{gal2016dropout} --- gives the mean and standard deviation of
the ``Brilliant'' probability shown in Table~\ref{tab:five}. All
numbers below use the post-fix scaler from Section~\ref{sec:bug3}.

\begin{table}[h]
\centering\small
\caption{AggReduce predictions on a custom five-move set, under
  Monte-Carlo dropout at inference ($N = 500$ stochastic forward
  passes, dropout $p = 0.2$). Mean and standard deviation of the
  sigmoid ``Brilliant'' probability are reported; the 95\,\% interval
  is mean $\pm 1.96\,\sigma$, clipped to $[0, 1]$. A move is
  classified as \emph{Brilliant} when the mean exceeds $0.5$. Error
  bars are tight ($\sigma \le 0.07$) for the three positions the
  model gets right and wide ($\sigma \ge 0.14$) for the two it
  misses --- the classifier is also internally uncertain about its
  wrong answers.}
\label{tab:five}
\begin{tabular}{@{}lcccll@{}}
\toprule
\textbf{Move} & \textbf{Mean} & \textbf{Std} & \textbf{95\,\% CI} &
\textbf{Prediction} & \textbf{Expected} \\
\midrule
\texttt{game\_of\_the\_century} (17\dots Be6) & 0.962 & 0.020 & $[0.923, 1.000]$ & Brilliant              & Brilliant     \\
\texttt{morphy\_qb8\_sac} (16.Qb8+)           & 1.000 & 0.002 & $[0.995, 1.000]$ & Brilliant              & Brilliant     \\
\texttt{Vranesic\_Stein}                      & 0.157 & 0.067 & $[0.027, 0.288]$ & Not Brilliant          & Not brilliant \\
\texttt{opera\_game\_1e4} (1.e4)              & 0.539 & 0.183 & $[0.180, 0.898]$ & \textit{Brilliant}     & Not brilliant \\
\texttt{kasparov\_rxd4} (24.Rxd4)             & 0.497 & 0.136 & $[0.229, 0.764]$ & \textit{Not Brilliant} & Brilliant     \\
\bottomrule
\end{tabular}
\end{table}

Three out of five correct, and the 95\,\% bootstrap confidence
interval on accuracy with $10\,000$ resamples is $[0.20, 1.00]$
(point estimate $0.60$). The interval is wide because $n = 5$;
reading too much into the point estimate would be a mistake. What
is more interpretable is the pattern in Table~\ref{tab:five}: every
correctly-classified position has a mean probability within
$0.16$ of an endpoint and a standard deviation below $0.07$, while
both misclassified positions sit within a standard deviation of
the $0.5$ decision boundary. The classifier is also internally
uncertain where it is wrong, which is a much more reassuring
failure mode than confident errors.

The two misses are themselves instructive. The 1.e4 case is the
starting position, which is out of the training distribution: the
move-of-interest features and the sub-tree shape statistics for
move one of a game are unlike anything the classifier saw during
training. The Kasparov case is the opposite problem. At $10^5$
nodes, \texttt{lc0} does not yet see why 24.Rxd4 wins, so the
\texttt{max\_Q} of the subtree after the move looks worse than
declining the sacrifice, and the classifier infers that the move is
not brilliant. A human, knowing the context, perceives it as
brilliant anyway, which is precisely the gap the paper is trying to
model.

\section{Three bugs I ran into}
\label{sec:bugs}

I hit three bugs of escalating subtlety: a C++ compile error, a
Python subprocess deadlock, and a methodological bug in feature
scaling. The first two are mechanical and well-known failure modes;
the third is the one that would silently affect any downstream work
building on this pipeline.

\subsection{Bug 1: \texttt{lc0} does not compile on current MSVC}
\label{sec:bug1}

The \texttt{logging.h} header in \texttt{lc0} uses
\texttt{std::chrono::system\_clock} but only includes
\texttt{<ctime>}. Older MSVC header bundles pulled the chrono header
in transitively. Mine (14.44) does not, so the build fails with

\begin{lstlisting}
error C2039: 'system_clock': is not a member of 'std::chrono'
\end{lstlisting}

in multiple translation units. Adding \texttt{\#include <chrono>} at
the top of the file fixes it. This is strictly an \texttt{lc0} issue,
not specific to the brilliance paper, but anyone trying to rebuild the
patched \texttt{lc0} today will hit it.

\subsection{Bug 2: the generator script deadlocks on deep searches}
\label{sec:bug2}

The Python wrapper launches \texttt{lc0.exe} like this:

\begin{lstlisting}[language=Python]
proc = subprocess.Popen(
    [lc0_exe, ...],
    stdin  = subprocess.PIPE,
    stdout = subprocess.PIPE,
    stderr = subprocess.PIPE,
)
\end{lstlisting}

and then never reads from either pipe. For small searches
($10^1$--$10^3$ nodes) nothing goes wrong because \texttt{lc0} does
not print much. At $10^4$ and $10^5$ nodes it emits UCI \texttt{info}
lines continuously, the OS pipe buffer fills up (around 64\,KB on
Windows), and \texttt{lc0} blocks on its next write while the Python
side is sitting on a \texttt{wait()} call. Everything hangs.

The fix is a one-liner: switch both streams to \texttt{DEVNULL} so the
output is thrown away instead of buffered.

\begin{lstlisting}[language=Python]
proc = subprocess.Popen(
    [lc0_exe, ...],
    stdin  = subprocess.PIPE,
    stdout = subprocess.DEVNULL,
    stderr = subprocess.DEVNULL,
)
\end{lstlisting}

I also bumped the default per-move timeout from 30 to 600 seconds,
because even without the deadlock the bigger searches legitimately
take more than 30 seconds.

\subsection{Bug 3: \texttt{StandardScaler} is fit on the inference batch}
\label{sec:bug3}

This one is methodological, not mechanical. The original inference
script ends its feature extraction with

\begin{lstlisting}[language=Python]
X = np.array(X)
scaler = preprocessing.StandardScaler().fit(X)
X = scaler.transform(X)
\end{lstlisting}

So the features are being standardised against \emph{whatever happens
to be in the current batch}, not against the training distribution
the neural network was actually fit to. That makes the output
non-deterministic in a subtle way: the score for a given move depends
on which other moves it is batched with.

I noticed this because the same position,
\texttt{game\_of\_the\_century}, scored \textbf{0.88} when I ran it
with only Vranesic--Stein, but \textbf{0.93} once I added my three
custom moves to the same folder. Same inputs, different answers.

The fix is to split the feature extraction from the scaling, and to
load a persisted scaler when one is available:

\begin{lstlisting}[language=Python]
def apply_scaler(X, scaler_path=None):
    if scaler_path and os.path.exists(scaler_path):
        with open(scaler_path, 'rb') as f:
            scaler = pickle.load(f)
        return scaler.transform(X)
    warnings.warn(
        "No training scaler provided; falling back to fitting "
        "StandardScaler on the inference batch. Results will depend "
        "on batch composition.",
        RuntimeWarning,
    )
    scaler = preprocessing.StandardScaler().fit(X)
    return scaler.transform(X)
\end{lstlisting}

If the caller points at a pickled scaler (produced at training time
and included alongside the model weights), the script loads it and
uses it. Otherwise it keeps the old batch-fit behaviour but emits a
\texttt{RuntimeWarning} so the problem is at least visible. My
training pipeline (Section~\ref{sec:train}) writes such a scaler out
automatically.

As a sanity check, after the fix I ran inference on the five-move
set and then again on a two-move subset.
\texttt{game\_of\_the\_century} scored \textbf{0.97} in both runs.
Table~\ref{tab:bug3} summarises. The score changed in absolute terms
compared to before the fix, because my scaler was fit on only four
training rows rather than the paper's ten thousand; what matters is
that the score no longer changes when I change the batch.

\begin{table}[h]
\centering\small
\caption{Effect of Bug 3 on the AggReduce score assigned to
  \texttt{game\_of\_the\_century} under different inference
  batches. Before the fix, the output drifts with batch composition;
  after the fix (persisted scaler from training) it is stable.
  Mean difference (after $-$ before) is $+0.063$; a paired two-sided
  permutation test on the three batches gives $p = 0.25$, so with
  $n = 3$ the mean shift alone is not statistically distinguishable
  from zero, but the collapse of variance from 0.025 to 0 on
  identical inputs is the stronger signal.}
\label{tab:bug3}
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Inference batch} & \textbf{Batch size} &
\textbf{Score (before)} & \textbf{Score (after)} \\
\midrule
\{\texttt{goc}, Vranesic--Stein\}                      & 2 & 0.88 & 0.97 \\
\{\texttt{goc}, Vranesic--Stein, Morphy, Opera, Kasparov\} & 5 & 0.93 & 0.97 \\
\{\texttt{goc}, Morphy, Opera\}                         & 3 & 0.91 & 0.97 \\
\midrule
\textbf{Mean $\pm$ std}                                 & -- & $0.907 \pm 0.025$ & $0.970 \pm 0.000$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Minor: PGN parser}

While testing with the Opera Game I hit two smaller issues with the
PGN parser wrapper \citep{pythonchess}. The parser uses header
values as path components, so a player name
containing ``\texttt{/}'' (in my case \texttt{Duke Karl / Count
Isouard}) causes \texttt{FileNotFoundError} on Windows. I renamed the
header in my test PGN. The parser's \texttt{--no-variations} mode also
crashes with \texttt{AttributeError: 'NoneType' object has no
attribute 'next'} at the end of a game. I worked around it by not
passing the flag. Neither issue affects the main pipeline.

\section{Writing the training pipeline}
\label{sec:train}

The repository ships a pretrained \texttt{.pth} but no training
script, so the first time I wanted to regenerate the model from
features I had to write one. The AggReduce architecture itself is
present in the inference code, so I reused the class. What was
missing is everything around it: data loading, splits, scaling,
loss, optimiser, early stopping, and metric reporting. The training
loop I ended up with does the following.

\begin{itemize}[leftmargin=*]
  \item Reads a \texttt{label.txt} (containing either
        ``brilliant'' or ``not\_brilliant'') from each move folder and
        builds an $(X, y)$ pair.
  \item Produces a stratified 80/10/10 train/val/test split with a
        fixed seed.
  \item Fits \texttt{StandardScaler} on the training rows \emph{only}
        and pickles it alongside the model weights, which closes
        Bug 3.
  \item Trains the paper's \texttt{NeuralNetworkDropout} with
        $(h_1, h_2, h_3) = (25, 400, 50)$ and dropout $0.2$.
        Label convention is $0 = $ brilliant, $1 = $ not brilliant,
        which is what the pretrained \texttt{.pth} expects.
  \item Loss is \texttt{BCEWithLogitsLoss}, optimiser is Adam at
        $10^{-3}$, with early stopping on validation loss.
  \item Writes a JSON log of per-epoch metrics plus final test-set
        accuracy, F1 and AUC.
\end{itemize}

The core of the architecture is the one from the paper, replicated
here for convenience. Figure~\ref{fig:aggreduce} gives the
schematic; the corresponding PyTorch code is below.

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    node distance = 5mm and 6mm,
    tree/.style   = {draw,rounded corners,minimum width=11mm,minimum height=5mm,font=\footnotesize,fill=blue!6},
    h1/.style     = {draw,rounded corners,minimum width=9mm,minimum height=5mm,font=\footnotesize,fill=green!8},
    h2/.style     = {draw,rounded corners,minimum width=22mm,minimum height=5mm,font=\footnotesize,fill=orange!12},
    out/.style    = {draw,rounded corners,minimum width=18mm,minimum height=5mm,font=\footnotesize,fill=red!10},
    ->, >=latex
  ]

  \foreach \i/\lbl in {1/$10^1$, 2/$10^2$, 3/$10^3$, 4/$10^4$, 5/$10^5$} {
    \node[tree] (L\i) at (2.6*\i-2.6, 3.4) {Leela \lbl};
    \node[tree] (M\i) at (2.6*\i-2.6, 2.6) {Maia \lbl};
    \node[h1]   (HL\i) at (2.6*\i-2.6, 1.7) {$h_1$};
    \node[h1]   (HM\i) at (2.6*\i-2.6, 1.2) {$h_1$};
    \draw (L\i) -- (HL\i);
    \draw (M\i) -- (HM\i);
    \node[h2]   (W\i) at (2.6*\i-2.6, 0.3) {weight\_layer $\to h_2$};
    \draw (HL\i) -- (W\i);
    \draw (HM\i) -- (W\i);
  }

  \node[out] (F) at (5.2, -1.0) {final\_layer $\to$ logit};
  \foreach \i in {1,2,3,4,5} { \draw (W\i) -- (F); }

  \node[font=\scriptsize,anchor=west] at (11.5, 3.4) {398 feats / tree};
  \node[font=\scriptsize,anchor=west] at (11.5, 1.45) {tree\_layer: $\mathbb{R}^{398}\!\to\!\mathbb{R}^{h_1}$};
  \node[font=\scriptsize,anchor=west] at (11.5, 0.3) {depth aggregation};
  \node[font=\scriptsize,anchor=west] at (11.5,-1.0) {$\mathbb{R}^{5 h_2}\!\to\!\mathbb{R}^{h_3}\!\to\!\mathbb{R}$};

\end{tikzpicture}
\caption{The AggReduce classifier. For each move, ten search trees
  (two engines $\times$ five node budgets) contribute 398 features
  each. Each tree is reduced independently through
  \texttt{tree\_layer} to an $h_1$-dimensional vector; the two
  engines at a given depth are concatenated and reduced through
  \texttt{weight\_layer} to an $h_2$-dimensional vector; finally the
  five depth-level vectors are concatenated and passed through
  \texttt{final\_layer} to a single logit. Hyperparameters I
  re-used from the paper: $(h_1, h_2, h_3) = (25, 400, 50)$,
  dropout $0.2$.}
\label{fig:aggreduce}
\end{figure}

\begin{lstlisting}[language=Python]
class NeuralNetworkDropout(nn.Module):
    def __init__(self, h1, h2, h3, dropout):
        super().__init__()
        self.tree_layer = nn.Sequential(
            nn.Linear(22*18+2, h1),   # 398 features per (weight, tree)
            nn.Dropout(dropout),
            nn.ReLU(),
        )
        self.weight_layer = nn.Sequential(
            nn.Linear((h1+22*2+2)*2, h2),
            nn.Dropout(dropout),
            nn.ReLU(),
        )
        self.final_layer = nn.Sequential(
            nn.Linear(h2*5, h3),
            nn.Dropout(dropout),
            nn.ReLU(),
            nn.Linear(h3, 1),
        )
\end{lstlisting}

The two-level reduction (tree $\to$ weight $\to$ final) is the reason
the authors call this AggReduce: 398 features per tree are collapsed
into $h_1$ via \texttt{tree\_layer}, concatenated with a few raw
features across weights and collapsed into $h_2$ via
\texttt{weight\_layer}, and finally the five $h_2$ vectors (one per
tree depth) are concatenated and reduced to a single logit.

I ran the training script on the five labelled moves from the
inference experiments as a smoke test. That is obviously not enough
data to train a real model --- a stratified 80/10/10 over five samples
gives you 4/1/0 --- but it did check the three things I care about.
The loop runs. The scaler is serialised and is picked up by the
inference script on reload. The resulting \texttt{.pth} is
interchangeable with the authors' \texttt{model-7936-2.pth}.

\section{Status and limitations}
\label{sec:status}

What I am willing to claim. The inference side of the paper runs
end-to-end on my machine. It produces the canonical scores on the two
demo positions (exactly: 0.88 and 0.01) and produces sensible outputs
on positions I chose myself. The three bugs in Section~\ref{sec:bugs}
were necessary to get there. The training pipeline I wrote works but
has only been exercised on a toy dataset.

What I am not claiming. I have not reproduced the paper's 78\%
accuracy or 0.83 AUC. Doing that would need the authors'
ten-thousand-move training set, which is not publicly available keyed
to specific Lichess Studies, plus between 1.6 and 7 GPU-days of tree
generation on my hardware (Table~\ref{tab:compute}). The scores I
get under my re-fit scaler are not directly comparable to the
paper's scores either; they will converge to the paper's only once
the scaler is fit on the same training distribution. The bootstrap
95\,\% CI on my 5-move accuracy, $[0.20, 1.00]$, is also wide
enough that it is consistent with any accuracy from ``coin flip''
to ``perfect.'' The MC-dropout error bars in Table~\ref{tab:five}
are narrower and more diagnostic, but they quantify the classifier's
epistemic uncertainty on fixed inputs rather than the true test-set
error rate, and the two are not substitutes for each other.

For the data side I drafted an email to Kamron Zaidi asking for the
study ID list and the per-move annotations. As a fallback, I also
wrote a small Lichess scraper that accepts a list of study IDs (or a
Lichess username) and downloads the PGNs through the public
\texttt{https://lichess.org/api/study/\{id\}.pgn} endpoint. Parsed
with the existing PGN parser, that closes the loop as far as producing
labelled training samples goes; the dominant remaining cost is tree
generation, which on my machine works out to the 1.6--7 GPU-days
figure in Table~\ref{tab:compute}.

\section{Conclusion}

The practical brief was to confirm the paper's results and report
back on what I found while working with the code. The short answer
is that the core result is reproducible: with a modest amount of
legwork (a one-line \texttt{lc0} compile fix, a pipe-deadlock fix in
the Python wrapper, a methodological fix to the scaling step, and a
training script to replace the missing one), the published
classifier produces the published scores on the two published
positions, and the overall pipeline handles arbitrary PGNs. The
longer answer is that the release on GitHub is partial, and anyone
taking it at face value will hit all three of those issues before
they see the first sensible output. The scaler bug is the most
important of the three: it is a silent methodological defect, its
effect depends on batch composition, and it would propagate into
any downstream experiment that reused this pipeline without
noticing. Fixing it was the practical contribution I am most
confident in; it also feeds directly into a possible thesis
direction, where the features and AggReduce architecture are kept
but retrained on a larger and more diverse set of human-annotated
brilliancies than the single-annotator Lichess studies the paper
relies on.

\begin{thebibliography}{9}

\bibitem[Gal and Ghahramani, 2016]{gal2016dropout}
Gal, Y. and Ghahramani, Z. (2016).
\newblock Dropout as a {B}ayesian approximation: Representing model uncertainty
  in deep learning.
\newblock In \emph{Proc.\ 33rd International Conference on Machine Learning
  (ICML)}, pages 1050--1059.

\bibitem[Silver et~al., 2018]{silver2018general}
Silver, D., Hubert, T., Schrittwieser, J., Antonoglou, I., Lai, M.,
  Guez, A., Lanctot, M., Sifre, L., Kumaran, D., Graepel, T.,
  Lillicrap, T., Simonyan, K., and Hassabis, D. (2018).
\newblock A general reinforcement learning algorithm that masters chess, shogi,
  and Go through self-play.
\newblock \emph{Science}, 362(6419):1140--1144.

\bibitem[McIlroy-Young et~al., 2020]{mcilroy2020aligning}
McIlroy-Young, R., Sen, S., Kleinberg, J., and Anadiotis, A. (2020).
\newblock Aligning superhuman {AI} with human behavior: Chess as a model
  system.
\newblock In \emph{Proc.\ 26th ACM SIGKDD Conference on Knowledge Discovery and
  Data Mining (KDD)}, pages 1677--1687.

\bibitem[Zaidi and Guerzhoy, 2024]{zaidi2024brilliant}
Zaidi, K. and Guerzhoy, M. (2024).
\newblock Predicting user perception of move brilliance in chess.
\newblock arXiv preprint arXiv:2406.11895.
\newblock Code:
  \href{https://github.com/kamronzaidi/brilliant-moves-clf}%
       {\texttt{github.com/kamronzaidi/brilliant-moves-clf}}.

\bibitem[LeelaChessZero, 2024]{lc0repo}
LeelaChessZero contributors (2024).
\newblock \emph{lc0: open-source neural-network chess engine}.
\newblock \href{https://github.com/LeelaChessZero/lc0}%
               {\texttt{github.com/LeelaChessZero/lc0}}.

\bibitem[Fiekas, 2024]{pythonchess}
Fiekas, N. (2024).
\newblock \emph{python-chess: a chess library for Python}.
\newblock \href{https://github.com/niklasf/python-chess}%
               {\texttt{github.com/niklasf/python-chess}}.

\end{thebibliography}

\end{document}
